Appendix B: Supplemental Results

### 2020-02-13 ms/supp/App_1.Rmd
updated appendixes

\section{General model structure}
\subsection{Integral Projection Models (IPMs)}
To generate fully-known true distributions for the virtual species, we used the general structure of an Integral Projection Model (IPM) to calculate the intrinsic growth rate $\lambda$ in each cell of the gridded landscape, and adapted the regression-based structure of an IPM into an individual-level, simulation-based cellular automata (CA) model to produce spatiotemporally dynamic abundance distributions.

@@

\newpage
\section{Regression equations for the virtual species}
We used the above IPM structure, including a seed bank and dispersal, as the basis for the virtual species. Thus, the population in each cell $i$ has size distribution $z$, where $z^{\prime}$ is the size distribution the next year, and $\boldsymbol{z_i}$ is a matrix with columns of 1 (intercept) and $z$. For the IPMs, we used the following regression equations to populate discretized transition matrices for each cell to then calculate the deterministic intrinsic growth rate $\lambda$. For the CA\textsubscript{i} models, we used the same regression equations to generate expected values for each individual, and then drew simulated outcomes from the specified distribtutions.

>>>>

\section{Integral Projection Model structure}
\subsection{Overview}
To generate fully-known true distributions for the virtual species, we used the general structure of an Integral Projection Model (IPM) to calculate the intrinsic growth rate $\lambda$ in each cell of the gridded landscape, and adapted the regression-based structure of an IPM into an individual-level, simulation-based cellular automata (CA) model to produce spatiotemporally dynamic abundance distributions.

@@

\newpage
\section{Virtual species simulation}
We used the above IPM structure, including a seed bank and dispersal, as the basis for the virtual species. Thus, the population in each cell $i$ has size distribution $z$, where $z^{\prime}$ is the size distribution the next year, and $\boldsymbol{z_i}$ is a matrix with columns of 1 (intercept) and $z$. For the IPMs, we used the following regression equations to populate discretized transition matrices for each cell to then calculate the deterministic intrinsic growth rate $\lambda$. For the CA\textsubscript{i} models, we used the same regression equations to generate expected values for each individual, and then drew simulated outcomes from the specified distribtutions.

### 2020-02-13 ms/supp/App_2.Rmd
updated appendixes

\newpage
\section{Sensitivity and specificity}

\subsection{Overall averages}
The True Skill Statistic (TSS) evaluates the combined ability of a model to predict the presences and the absences. While TSS is useful for comparing overall performance, the individual components can provide insight as well. Here, we show summary plots for the effect of each scenario on the sensitivity (the proportion of true presences correctly predicted as presences) and specificity (the proportion of true absences correctly predicted as absences).

@@

\newpage
\subsection{Scenario effects}
Some scenarios, such as \emph{non-equilibrium} for MaxEnt or \emph{under-dispersal} for the process-based models, result in consistent under-prediction, as seen by a decline in sensitivity and an increase in specificity. In contrast, \emph{incorrect covariates} has more complex effects, leading to a decline in both sensitivity and specificity in the process-based models. In this case, not only are presences more poorly predicted, but absences are more poorly predicted as well.

\begin{figure}
	\centering\includegraphics[width=.9\linewidth]{../../figs/Supp_SensvIdeal.jpg}
	\caption{\label{fig:SensitivityvIdeal} Effect of scenario on median sensitivity relative to the \emph{ideal} scenario. Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
\end{figure}

@@

\newpage
\section{Estimated slopes}
The IPM and CA\textsubscript{i} models fit regressions of the same form as the generative models, and so error in the estimated slopes can be directly compared. The pattern of accuracy across scenarios reflects the observed trends in TSS. The effect of data scenario is minimal, with the exception of non-equilibrium for the shrub. The slopes clearly show the improvement from sampling newer populations with a broader age distribution, relative to all other scenarios where populations tended to be older. This effect is absent from the biennial regressions, as individuals are not long-lived.
\begin{figure}
	\centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
	\caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the vital rate regressions for the shrub. Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models. The true value is indicated by the vertical dashed line. For (a) survival probability, the slopes are estimated more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution. In contrast, (b) growth, (c) flowering probability, and (e) germination probability show little variation among scenarios, while for (d) seed production, added \emph{measurement error} increased variability. Note that the axes vary among panels.}
\end{figure}

>>>>

\section{Sensitivity and specificity}

\subsection{Overall averages}
The True Skill Statistic (TSS) evaluates the combined ability of a model to predict the presences and the absences. While TSS is useful for comparing overall performance, the individual components can provide insight as well. Here, we show summary plots for the effect of each scenario on the sensitivity (the proportion of true presences correctly predicted as presences) and specificity (the proportion of true absences correctly predicted as absences). 
\clearpage

@@

\clearpage
\subsection{Scenario effects}
Some scenarios, such as \emph{non-equilibrium} for MaxEnt or \emph{under-dispersal} for the process-based models, result in consistent under-prediction, as seen by a decline in sensitivity and an increase in specificity. In contrast, \emph{incorrect covariates} has more complex effects, leading to a decline in both sensitivity and specificity in the process-based models. In this case, not only are presences more poorly predicted, but absences are more poorly predicted as well.

\clearpage
\begin{figure}
	\centering\includegraphics[width=.9\linewidth]{../../figs/Supp_SensvIdeal.jpg}
	\caption{\label{fig:SensitivityvIdeal} Effect of scenario on median sensitivity relative to the \emph{ideal} scenario. Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
\end{figure}

@@

\clearpage
\section{Estimated slopes}
The IPM and CA\textsubscript{i} models fit regressions of the same form as the generative models, and so error in the estimated slopes can be directly compared. The pattern of accuracy across scenarios reflects the observed trends in TSS. The effect of data scenario is minimal, with the exception of \emph{non-equilibrium} for the shrub survival probability regressions. The slopes clearly show the improvement from sampling newer populations with a broader age distribution, relative to all other scenarios where populations tended to be older. This effect is absent from the biennial regressions, as individuals are not long-lived. Additionally, slopes for the growth regression were poorly estimated for the shrub, resulting from the imposition of a size maximum in the virtual species.

\begin{figure}
	\centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
	\caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the vital rate regressions for the shrub. Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models. The true value is indicated by the vertical dashed line. For (a) survival probability, the slopes are estimated more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution. In contrast, (b) growth, (c) flowering probability, and (e) germination probability show little variation among scenarios, while for (d) seed production, added \emph{measurement error} increased variability. Note that the axes vary among panels.}
\end{figure}

### 2020-02-13 ms/supp/App_B.tex
updated appendixes

\newpage
\section{Sensitivity and specificity}

@@

The True Skill Statistic (TSS) evaluates the combined ability of a model
to predict the presences and the absences. While TSS is useful for
comparing overall performance, the individual components can provide
insight as well. Here, we show summary plots for the effect of each
scenario on the sensitivity (the proportion of true presences correctly
predicted as presences) and specificity (the proportion of true absences
correctly predicted as absences).

@@

\newpage
\subsection{Scenario effects}

@@

\begin{figure}
    \centering\includegraphics[width=.9\linewidth]{../../figs/Supp_SensvIdeal.jpg}
    \caption{\label{fig:SensitivityvIdeal} Effect of scenario on median sensitivity relative to the \emph{ideal} scenario. Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
\end{figure}

@@

\newpage
\section{Estimated slopes}

The IPM and CA\textsubscript{i} models fit regressions of the same form
as the generative models, and so error in the estimated slopes can be
directly compared. The pattern of accuracy across scenarios reflects the
observed trends in TSS. The effect of data scenario is minimal, with the
exception of non-equilibrium for the shrub. The slopes clearly show the
improvement from sampling newer populations with a broader age
distribution, relative to all other scenarios where populations tended
to be older. This effect is absent from the biennial regressions, as
individuals are not long-lived.

\begin{figure}
    \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
    \caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the shrub. Survival regressions for the shrub are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution. Note that the axes vary among panels.}
\end{figure}

\begin{figure}
    \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_biennial.pdf}
    \caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios. Peaks at 0 represent data sets where the optimal model did not include that term. Note that the axes vary among panels.}
\end{figure}

>>>>

\section{Sensitivity and specificity}

@@

The True Skill Statistic (TSS) evaluates the combined ability of a model
to predict the presences and the absences. While TSS is useful for
comparing overall performance, the individual components can provide
insight as well. Here, we show summary plots for the effect of each
scenario on the sensitivity (the proportion of true presences correctly
predicted as presences) and specificity (the proportion of true absences
correctly predicted as absences). \clearpage

@@

\clearpage
\subsection{Scenario effects}

@@

\clearpage
\begin{figure}
    \centering\includegraphics[width=.9\linewidth]{../../figs/Supp_SensvIdeal.jpg}
    \caption{\label{fig:SensitivityvIdeal} Effect of scenario on median sensitivity relative to the \emph{ideal} scenario. Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
\end{figure}

@@

\clearpage
\section{Estimated slopes}

The IPM and CA\textsubscript{i} models fit regressions of the same form
as the generative models, and so error in the estimated slopes can be
directly compared. The pattern of accuracy across scenarios reflects the
observed trends in TSS. The effect of data scenario is minimal, with the
exception of \emph{non-equilibrium} for the shrub survival probability
regressions. The slopes clearly show the improvement from sampling newer
populations with a broader age distribution, relative to all other
scenarios where populations tended to be older. This effect is absent
from the biennial regressions, as individuals are not long-lived.
Additionally, slopes for the growth regression were poorly estimated for
the shrub, resulting from the imposition of a size maximum in the
virtual species.

\begin{figure}
    \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
    \caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the vital rate regressions for the shrub. Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models. The true value is indicated by the vertical dashed line. For (a) survival probability, the slopes are estimated more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution. In contrast, (b) growth, (c) flowering probability, and (e) germination probability show little variation among scenarios, while for (d) seed production, added \emph{measurement error} increased variability. Note that the axes vary among panels.}
\end{figure}

\begin{figure}
    \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_biennial.pdf}
    \caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the vital rate regressions for the biennial. Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models. The true value is indicated by the vertical dashed line. The fitted slopes were consistent with minimal differences among data scenarios for (a) survival probability, (b) growth, (c) flowering probability, (d) seed production, and (e) germination probability. Peaks at 0 represent data sets where the optimal model did not include that term. Note that the axes vary among panels.}
\end{figure}
